Embedded Scalars: {A}n Experimental Revisit

### 2013-01-23 text/paper/paper3.tex
conclusions sec com

\title{Scalar Items in Embedded Position: {A}n Experimental Revisit}
\author{Fabian Schlotterbeck, Michael Franke and Petra Augurzky}
\date{}

@@

\begin{abstract}
  Scalar item \emph{some} is widely assumed to readily undergo a
  pragmatic meaning enrichment to \emph{some but not all} if it occurs
  in matrix position. The question under which circumstances this
  meaning enrichment can take place in embedded position is an
  important one to decide the delineation between semantics and
  pragmatics, and therefore currently under renewed debate. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present a novel experimental paradigm to test whether so-called
  local readings are available with or without prosodic markedness. On
  top of adding prosodic manipulation, our experimental design was
  catered to a number of methodological difficulties that were
  identified for previous studies, such as the possibility of
  confounds of pictorial properties. The data obtained refutes both a
  traditional Gricean view of scalar implicature calculation, as well
  as the more recent grammatical approach.
\end{abstract}

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism about availability
and preference of readings of our target sentences. The intonational
markedness hypothesis, which is frequently relied on to back-up
traditionalism, is elaborated on in
Section~\ref{sec:inton-mark-hypoth}, together with a concise
assessment of the sparse empirical data on the effects of intonation
on the interpretation of (embedded) scalar
items. Section~\ref{sec:previous-studies} recaps the results of
previous related studies, pointing out the most important necessary
amendments that motivated out own design. Section~\ref{sec:exp}
eventually 
describes our experimental design and continues to state, analyze and interpret the
results we obtained. We defend our design and the interpretation of
the results based on it in Section~\ref{sec:conclusions}.

@@

Finally, we should address the obvious worry that maybe our results
could have been the product of typicality after all. Remember that
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue that the
distribution of responses to \as-sentences found in previous studies
could in part be explained by differences in typicality of the
pictures with respect to the possible readings. In the present case
this kind of reasoning would have to be extended to either cover the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty. On intuitive grounds,
we take either extension to be far-fetched and implausible. But even
if this line of explanation can be plausibly extended to incrementally
revealed picture verification, this does not as such save the day for
traditionalism or grammaticalism, for the same reasons as given above.

@@

>>>>

\title{Embedded Scalars: {A}n Experimental Revisit} \author{Fabian
  Schlotterbeck, Michael Franke and Petra Augurzky} \date{}

@@

\begin{abstract}
  Scalar item \emph{some} is widely assumed to undergo, given relevant
  preconditions, a pragmatic meaning enrichment to \emph{some but not
    all} if it occurs in matrix position. The question under which
  circumstances this meaning enrichment can take place in embedded
  position plays an important role in deciding how to delineate
  semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present a novel experimental paradigm to test whether the
  relevant pragmatic enrichment is available in embedded positions
  with or without prosodic markedness. On top of adding prosodic
  manipulation, our experimental design was catered to a number of
  methodological difficulties that were identified for previous
  related studies. The data obtained refutes both a traditional
  Gricean view of scalar implicature calculation, as well as the more
  recent grammatical approach.
\end{abstract}

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism about availability
and preference of readings of our target sentences. The intonational
markedness hypothesis, which is frequently relied on to back-up
traditionalism, is elaborated on in
Section~\ref{sec:inton-mark-hypoth}, together with a concise
assessment of the sparse empirical data on the effects of intonation
on the interpretation of (embedded) scalar
items. Section~\ref{sec:previous-studies} recaps the results of
related experimental studies, pointing out the most important
necessary amendments that motivated out own
design. Section~\ref{sec:exp} eventually describes our experimental
design and continues to state, analyze and interpret the results we
obtained.

@@

Finally, we should address the obvious worry that maybe our results
could have been the product of typicality after all. Remember that
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue that the
distribution of responses to \as-sentences found in previous studies
could be explained by differences in typicality, given an
\as-sentence, of the pictures used by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev}. In the present case this
kind of reasoning would have to be extended to either cover the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty. On intuitive grounds,
we take either extension to be far-fetched and implausible. But even
if this line of explanation can be plausibly extended to incrementally
revealed picture verification, this does not as such save the day for
traditionalism or grammaticalism, because, as argued above, typicality
would likely still not predict the change in categorical truth-value
judgement that we found in our experiment.

@@

\paragraph{Future Directions.} The data obtained from our experiment
raises a number of interesting questions for future research. Firstly,
we found a surprisingly strong tendency of subjects to give literal or
local answers rather consistently. This seems to be related to what
has been reported in a number of previous studies on unembedded
\emph{some}, where a tendency was found as well in subjects to either
consistently interpret \emph{some} `semantically' as \emph{some and
  possibly all} or to consistently interpret it `pragmatically' as
\emph{some and not all}
\citep[e.g.][]{NoveckPosada2003:Characterizing-,BottNoveck2004:Some-Utterances,DegenTanenhaus2012:Processing-Scal}. This
inter-subjective consistency has not been paid due attention to in
previous studies on embedded scalar items. It would be interesting to
test whether there is any indication of a correlation between, on the
one hand, subjects' tendencies to give literal vs.~local answers for
embedded occurrences of \emph{some} and, on the other hand, subjects'
tendencies to interpret unembedded \emph{some} semantically or
pragmatically. If such a correlation exists, it would point towards
the possibility that different subjects may even have different
lexical representations of \emph{some} or possibly employ different
processes of interpreting it. If so, this would severely mix up the
theoretical landscape and should also be taken into account in any
further study on the processing of scalar items.
